---
title: "RMA"
output: html_document
date: "2023-11-11"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, }
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse, 
               knitr,
               here,
               DT,
               readxl, 
               metafor,
               clubSandwich,
               orchaRd, 
               MuMIn,
               patchwork,
               GoodmanKruskal,
               networkD3,
               ggplot2,
               plotly,
               ggsignif,
               visdat,
               ggalluvial,
               ggthemr, 
               cowplot,
               grDevices,
               png,
               grid,
               gridGraphics,
               pander,
               formatR,
               rmdformats,
               pander,
               metapower, 
               cowplot, 
               gt, 
               ggpubr)

library(metameta)
```

#### SIMPLE INTERVENTIONS ####

# Read in data 

```{r}
SI <- dataall_LSR2 %>% 
  filter(CohortType == "Exercise intervention")
```

## Label outcome types

```{r}
SI <- SI %>% 
  mutate(OutcomeType = case_when(str_detect(OutcomeLabel, regex("lma|locomotor|oft|horizontal|distance|vertical|climbing", ignore_case = TRUE)) ~ "Locomotor activity",
                                 str_detect(OutcomeLabel, "PPI") ~ "Prepulse inhibition", 
                                 TRUE ~ `Type of outcome?`)) %>% 
  relocate(OutcomeType, .after = OutcomeLabel)

SI %>% group_by(OutcomeType, OutcomeLabel, `Type of outcome?`) %>% count()
```


##################################################################################################### Calculate effect sizes #####################################################################################################

## Join ErrorType from LSR3SYRF original download onto the SI dataframe to label which error measurement used in each study 
Checked in excel that only one type of error measurement is used per study. 

```{r}
LSR3_SRYF101123_studyerrortype <- LSR3_TAAR_agonists_for_psychosis_101123_7 %>% 
  select(StudyId, OutcomeId, ErrorType) %>% 
  distinct()

SI <- left_join(SI, LSR3_SRYF101123_studyerrortype, by = c("StudyId", "OutcomeId"))
```


## 1. Calculate SD for all comparisons 
F = final
C/T/S = control/treatment/sham
L/n/m/v = cohort label/n in cohort/mean/variance

```{r}
SI <- SI %>% 
  mutate(F_C_v.SD = case_when(ErrorType == "IQR" ~ (F_C_v/1.35), 
                           ErrorType == "SD" ~ F_C_v, 
                           ErrorType == "SEM" ~ sqrt(F_C_n)*F_C_v)) %>%  
  relocate(F_C_v.SD, .after = F_C_v) %>% 
  mutate(F_T_v.SD = case_when(ErrorType == "IQR" ~ (F_T_v/1.35), 
                           ErrorType == "SD" ~ F_T_v, 
                           ErrorType == "SEM" ~ sqrt(F_T_n)*F_T_v)) %>%  
  relocate(F_T_v.SD, .after = F_T_v) %>% 
  mutate(F_S_v.SD = case_when(ErrorType == "IQR" ~ (F_S_v/1.35), 
                           ErrorType == "SD" ~ F_S_v, 
                           ErrorType == "SEM" ~ sqrt(F_S_n)*F_S_v)) %>%  
  relocate(F_S_v.SD, .after = F_S_v)
 
```

### Check how many comparisons will be able to calculate NMD for 

```{r}
SI <- SI %>%
  rowwise() %>%
  mutate(`NMD possible` = all(!is.na(F_S_L)))
```

## 2. Calculate effect sizes (SMD for all, NMD where possible)

### Calculate true n for control groups (n'c)
Since a single experiment can contain a number of comparisons, a control group can serve more than one treatment group. Were this control cohort to be included in more than one comparison, it would be represented more than once in the sum- mary estimates calculated. To avoid this, we recommend to correct the number of animals reported in the control group by dividing the reported number by the number of treatment groups served in order to give a “true number of control animals”. This corrected number can then be used when calculating the total number of animals in the meta-analysis and where the number of animals is used in the weighting of effect sizes

```{r}
F_C_L_frequencies <- SI %>%
  group_by(StudyId, OutcomeId, F_C_L) %>%
  summarise(Frequency_FCL = n()) %>% 
  ungroup()

# Step 2: Join the frequencies back to the original dataframe
SI <- SI %>%
  left_join(F_C_L_frequencies)

# Step 3: Divide F_C_n by the frequency count
SI <- SI %>%
  mutate(F_C_n_true = F_C_n / Frequency_FCL)

```

### SMD

Hedges g to account for small sample sizes (default for SMD when using the escalc() function) - Hedge’s g (statistically corrects for variance that may be introduced when sample sizes are small (Larry V. Hedges 1981))
m1i = control, m2i = rx 

```{r}
# Hedge's g effect size
SMD_data_SI.nottrue <- escalc(
  measure = "SMD", 
  m1i = SI$F_C_m, 
  m2i = SI$F_T_m, 
  sd1i = SI$F_C_v.SD, 
  sd2i = SI$F_T_v.SD, 
  n1i = SI$F_C_n, 
  n2i = SI$F_T_n, 
  data = SI) %>% 
  select(yi, vi, )

SI$SMD <- SMD_data_SI.nottrue$yi
SI$SMDv <- SMD_data_SI.nottrue$vi
  
SMD_data_SI.true<- escalc(
  measure = "SMD", 
  m1i = SI$F_C_m, 
  m2i = SI$F_T_m, 
  sd1i = SI$F_C_v.SD, 
  sd2i = SI$F_T_v.SD, 
  n1i = SI$F_C_n_true, 
  n2i = SI$F_T_n, 
  data = SI)


#escalc (m1 - m2) = (control - treatment)
# so if greater is better, then *-1

SI$SMD_true <- SMD_data_SI.true$yi #Row 112, SMD calculated, but SMD_true not calculated
SI$SMDv_true <- SMD_data_SI.true$vi

```

### NMD
Assume that treatments are closer to shams than controls are 
So C-S > T-S 
```{r}
# NMD
SI <- SI %>% 
  mutate(`NMD (%)` = 100*(((F_C_m - F_S_m) - (F_T_m - F_S_m))/(F_C_m - F_S_m))) %>% 
  mutate(`NMD_SDc*` = 100*((F_C_v.SD/(F_C_m - F_S_m)))) %>% 
  mutate(`NMD_SDrx*` = 100*((F_T_v.SD/(F_C_m - F_S_m)))) %>% 
  mutate(NMD_v = sqrt(((`NMD_SDc*`)^2/F_C_n_true) + ((`NMD_SDrx*`)^2/F_T_n))) 

SI.direction <- SI %>% 
  mutate(SMD = if_else((GreaterIsWorse = FALSE), -1*SMD, SMD)) %>% 
  mutate(SMD_true = if_else((GreaterIsWorse = FALSE), -1*SMD_true, SMD_true)) %>% 
  rename(CategoryDiseaseInduction = `Category of disease model induction method:`)

diagnostic <- SI.direction %>% 
  select(F_C_m, F_T_m, F_S_m, SMD, SMD_true, `NMD (%)`, GreaterIsWorse) %>% 
  mutate(CbiggerT = if_else(F_C_m > F_T_m, "Yes", "No"))
```


## 3. Fit Random Effects Model

Random effects structure: 
StudyId - variance in effect sizes attributed to variance of effect sizes between studies


# By drug name SMD

```{r}
SI.direction.meta <- SI.direction %>% 
  left_join(LSR3_TAAR_agonists_for_psychosis_101123_7)


SI.SMDmeta <- rma.mv(yi = SMD, V = SMDv, 
                         random = ~1 | StudyId / `Animal strain?` / CategoryDiseaseInduction / OutcomeType, 
                         data = SI.direction.meta, 
                         method = 'REML')
summary(SI.SMDmeta)

SI.SMDmeta.drug <- rma.mv(yi = SMD, V = SMDv, 
                         random = ~1 | StudyId / `Animal strain?` / CategoryDiseaseInduction / OutcomeType, 
                         mods = ~ drugname1, 
                         data = SI.direction.meta, 
                         method = 'REML')

SI.SMDmeta.drug$sigma2

```

tau^2 estimates for SI.SMD.drug:
StudyId = 2.543926e-08
StudyId/Animal strain = 2.257189e-01
StudyId/Animal strain?/CategoryDiseaseInduction = 1.199150e-09
StudyId/Animal strain?/CategoryDiseaseInduction/OutcomeType = 6.827822e-01

pooled SMD estimate = 0.8472, ci.lb = 0.4540, ci.ub = 1.2404

```{r}
#SMD.meta.drug <- read_csv("~/Library/Mobile Documents/com~apple~CloudDocs/Desktop/PhD /LSRs/Meta-Analysis/RMA/SMD_meta_drug.csv")
pooled_SMD_estimate <- 0.8472
pooled_SMD_cilb <- 0.4540
pooled_SMD_ciub <- 1.2404

pooled_SMD_data <- data.frame(
  drugname1 = 'Pooled estimate',
  estimate = pooled_SMD_estimate,
  ci.lb = pooled_SMD_cilb,
  ci.ub = pooled_SMD_ciub
  )

SMD.meta.drug <- as.data.frame(SI.SMDmeta.drug[["b"]])


# Create a data frame for the pooled_SMD estimate
SMD.meta.drug <- as.data.frame(SMD.meta.drug)
SMD.meta.drug$drugname1 <- rownames((SMD.meta.drug))
SMD.meta.drug$ci_l <- SI.SMDmeta.drug[["ci.lb"]]
SMD.meta.drug$ci_u <- SI.SMDmeta.drug[["ci.ub"]]
rownames(SMD.meta.drug) <- seq_len(nrow(SMD.meta.drug))
SMD.meta.drug <- SMD.meta.drug[,c(2,1,3,4)]
colnames(SMD.meta.drug) <- colnames(pooled_SMD_data)
SMD.meta.drug$drugname1 <- as.character(SMD.meta.drug$drugname1)
SMD.meta.drug$drugname1 <- gsub("drugname1", "",SMD.meta.drug$drugname1)
SMD.meta.drug <- SMD.meta.drug[-1,]
SMD.meta.drug <- rbind(SMD.meta.drug, pooled_SMD_data)

SMD.meta.drug$facet_variable <- factor(SMD.meta.drug$facet_variable, levels = c("Compound 50A","Compound 50B","LK000764","RO5073012","RO5166017","RO5203648","RO5256390","RO5263397",  "SEP", "SEP-363856", "Pooled estimate"))
SMD.meta.drug$drugname1 <- factor(SMD.meta.drug$drugname1, levels = c("Compound 50A","Compound 50B","LK000764","RO5073012","RO5166017","RO5203648","RO5256390","RO5263397",  "SEP", "SEP-363856", "Pooled estimate"))
library(forcats)

# Reorder levels using forcats
SMD.meta.drug$drugname2 <- fct_relevel(SMD.meta.drug$drugname1, 
                                            "Compound 50A", "Compound 50B", "LK000764", "RO5073012", "RO5166017", 
                                            "RO5203648", "RO5256390", "RO5263397", "SEP", "SEP-363856", "Pooled estimate")


SMD.drug <- SMD.meta.drug %>% 
  group_by(drugname1) %>%
  ggplot(aes(x = "", y= estimate, ymin=ci.lb, ymax=ci.ub))+
  scale_x_discrete(limits=rev) +
  scale_y_continuous(name="Standardised Mean Difference (%)", limits=(c(-5, 5)))+
  expand_limits(y=c(-5, 5))+
  geom_pointrange() +
  geom_hline(yintercept =0, linetype=2)+
  geom_errorbar(aes(ymin=ci.lb, ymax=ci.ub),width=0.05, cex=0)+
  theme_minimal()+
  coord_flip()+
  facet_wrap(~ drugname2, strip.position="left", labeller = label_wrap_gen(width=20), nrow = 11, scales = "free_y")+
  geom_rect(aes(fill = drugname1),xmin = -Inf,xmax = Inf,
            ymin = -Inf,ymax = Inf,alpha = 0.1) +
  geom_text(aes(label = paste("95CI: [", ci.lb, ", ", ci.ub, "]")), 
            hjust = 1, vjust = -0.5,  # Adjust these values as needed for text placement
            size = 3, color = "black") +
  geom_point(shape=1, size=0.3) +
  #geom_point(shape=1, size=0.3) + 
  #geom_point(data = pooled_SMD_data, aes(x = "", y = estimate), shape = 18, size = 4, color = "red") +
  #geom_errorbar(data = pooled_SMD_data, aes(x = "", ymin = ci.lb, ymax = ci.ub), width=0.2, color = "red") + 
  #geom_text(data = pooled_SMD_data, aes(y = estimate, label = paste("95% CI: [", ci.lb, ", ", ci.ub, "]")), 
  #          hjust = 1, vjust = -0.7,  # Adjust hjust and vjust for correct positioning
  #          size = 3, color = "red") +
  labs(title = 'Subgroup analysis of TAAR1 Agonist Efficacy Using SMD (%)')+
  theme(plot.title=element_text(size=18,face="bold"),
        legend.position = "none",
        axis.text.y=element_text(),
        axis.ticks.y=element_blank(),
        axis.text.x=element_text(face="bold", ),
        axis.title.x =element_text(size=10,face="bold"),
        axis.title.y = element_blank(),
        strip.text.y.left = element_text(hjust=0,vjust = 0.5,angle=0,face="bold", size=10),
        strip.placement = "outside") 


SMD.drug

```

# By drug name NMD

```{r}
SI.NMDmeta <- rma.mv(yi = `NMD (%)`, V = NMD_v, 
                         random = ~1 | StudyId / `Animal strain?` / CategoryDiseaseInduction / OutcomeType, 
                         data = SI.direction.meta, 
                         method = 'REML')

drug_names_NMD <- SI.direction %>% filter(!is.na(`NMD (%)`)) %>% group_by(drugname1) %>% count()

SI.NMDmeta.drug <- rma.mv(yi = `NMD (%)`, V = NMD_v, 
                         random = ~1 | StudyId / `Animal strain?` / CategoryDiseaseInduction / OutcomeType, 
                         mods = ~ drugname1,
                         data = SI.direction.meta, 
                         method = 'REML')

SI.NMDmeta.drug$sigma2
```

tau^2 estimates for SI.NMD.drug: 
StudyId = 5.253489e+02
StudyId/Animal strain = 1.597956e-05
StudyId/Animal strain?/CategoryDiseaseInduction = 1.597957e-05
StudyId/Animal strain?/CategoryDiseaseInduction/OutcomeType = 2.827994e+02

Pooled estimate = 49.9161, ci.lb = 26.5976, ci.ub = 71.7525

```{r}
NMD.meta.drug <- read_csv("~/Library/Mobile Documents/com~apple~CloudDocs/Desktop/PhD /LSRs/Meta-Analysis/RMA/NMD.meta.drug.csv") %>% filter(!is.na(drugname1))

pooled_NMD_estimate <- 49.1751
pooled_NMD_cilb <- 26.5976
pooled_NMD_ciub <- 71.7525

# Create a data frame for the pooled_NMD estimate
pooled_NMD_data <- data.frame(
  drugname1 = "Pooled Estimate",
  estimate = pooled_NMD_estimate,
  ci.lb = pooled_NMD_cilb,
  ci.ub = pooled_NMD_ciub
) 

# Tau estimates
tau_info.NMD.drug <- paste(
  "Tau^2 Estimates:",
  "StudyId = 5.25e+02;",
  "StudyId/Animal strain = 1.60e-05;",
  "StudyId/Animal strain/CategoryDiseaseInduction = 1.60e-05;",
  "StudyId/Animal strain/CategoryDiseaseInduction/OutcomeType = 2.83e+02"
)

NMD.drug <- SMD.meta.drug %>% 
  group_by(drugname1) %>%
  ggplot(aes(x = "", y= estimate, ymin=ci.lb, ymax=ci.ub))+
  scale_x_discrete(limits=rev) +
  scale_y_continuous(name="Normalised Mean Difference (%)", limits=(c(-100, 100)))+
  expand_limits(y=c(-100, 100))+
  geom_pointrange() +
  geom_hline(yintercept =0, linetype=2)+
  geom_errorbar(aes(ymin=ci.lb, ymax=ci.ub),width=0.05, cex=0)+
  theme_minimal()+
  coord_flip()+
  facet_wrap(~ drugname1, strip.position="left", labeller = label_wrap_gen(width=20), nrow = 25, scales = "free_y")+
  geom_rect(aes(fill = drugname1),xmin = -Inf,xmax = Inf,
            ymin = -Inf,ymax = Inf,alpha = 0.1) +
  geom_text(aes(label = paste("95CI: [", ci.lb, ", ", ci.ub, "]")), 
            hjust = 1, vjust = -0.5,  # Adjust these values as needed for text placement
            size = 3, color = "black") +
  geom_point(shape=1, size=0.3)+
  geom_point(shape=1, size=0.3) + 
  geom_point(data = pooled_SMD_data, aes(x = "", y = estimate), shape = 18, size = 4, color = "red") +
  geom_errorbar(data = pooled_SMD_data, aes(x = "", ymin = ci.lb, ymax = ci.ub), width=0.2, color = "red") + 
  geom_text(data = pooled_SMD_data, aes(y = estimate, label = paste("95% CI: [", ci.lb, ", ", ci.ub, "]")), 
            hjust = 1, vjust = -0.7,  # Adjust hjust and vjust for correct positioning
            size = 3, color = "red") +
  labs(title = 'Subgroup analysis of TAAR1 Agonist Efficacy Using NMD (%)')+
  theme(plot.title=element_text(size=18,face="bold"),
        legend.position = "none",
        axis.text.y=element_text(),
        axis.ticks.y=element_blank(),
        axis.text.x=element_text(face="bold", ),
        axis.title.x =element_text(size=10,face="bold"),
        axis.title.y = element_blank(),
        strip.text.y.left = element_text(hjust=0,vjust = 0.5,angle=0,face="bold", size=10),
        strip.placement = "outside") 

NMD.drug
```
currently showing confidence intervals rather than prediction intervals - how do I get prediction intervals? 





